Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-26 paper/manuscript.tex
tweaks to merge H1 corr update

\subsection*{Measurements made on an H1 HA are less informative for understanding the evolution of H3 influenza}
To determine how broadly experimental measurements can be generalized across HAs, we repeated the foregoing analysis of H3N2 mutation frequencies, but using mutational effects measured in our prior deep mutational scanning of the WSN/1933 H1 HA~\cite{doud2016accurate} (Figure~\ref{suppfig:WSNprefs_logoplot}), which is highly diverged from the Perth/2009 H3 HA (the two HAs only have 42\% protein sequence identity).
Figure~\ref{fig:muteffect_maxfreq_WSN} shows that the correlations between the H1 experimental measurements and the maximum frequency that mutations reach during H3N2 viral evolution are consistently weaker than those using H3 experimental measurements (Figure~\ref{fig:muteffect_maxfreq}).
Therefore, the utility of an experiment for understanding natural evolution degrades as the experimental sequence becomes more diverged from the natural sequences that are being studied.

@@

\subsection*{There are large differences between H3 and H1 HAs in the amino-acid preferences of many sites}
An obvious hypothesis for why the H1 deep mutational scanning is not useful for understanding the evolution of H3N2 influenza viruses is that the effect of the same mutation is often different between these two HA subtypes.
To determine if this is the case, we examined how much the amino-acid preferences of homologous sites have shifted between H3 and H1 HAs.
Prior experiments have found only modest shifts in amino-acid preferences between two variants of influenza nucleoprotein with 94\% amino-acid identity~\cite{doud2015site} and variants of HIV envelope (Env) with 86\% amino-acid identity~\cite{haddox2018mapping}.
However, the H1 and H3 HAs are far more diverged, with only 42\% amino-acid identity (Figure~\ref{fig:distance_distribution}A).
One simple way to investigate the extent of shifts in amino-acid preferences is to correlate measurements from independent deep mutational scanning replicates on H1 and H3 HAs.
Figure~\ref{fig:distance_distribution}B shows that replicate measurements on the same HA variant are more correlated than those on different HA variants.

>>>>

\subsection*{Measurements made on an H1 HA are less informative for understanding the evolution of H3 influenza}
To determine how broadly experimental measurements can be generalized across HAs, we repeated the foregoing analysis of H3N2 mutation frequencies, but using mutational effects measured in our prior deep mutational scanning of the WSN/1933 H1 HA~\cite{doud2016accurate} (Figure~\ref{suppfig:WSNprefs_logoplot}), which is highly diverged from the Perth/2009 H3 HA (the two HAs only have 42\% protein sequence identity).
Figure~\ref{fig:muteffect_maxfreq_WSN} shows that the correlations between the H1 experimental measurements and the maximum frequency that mutations reach during H3N2 viral evolution are consistently weaker than those using H3 experimental measurements (compare Figure~\ref{fig:muteffect_maxfreq_WSN} to Figure~\ref{fig:muteffect_maxfreq}A).
Therefore, the utility of an experiment for understanding natural evolution degrades as the experimental sequence becomes more diverged from the natural sequences that are being studied.

@@

\subsection*{There are large differences between H3 and H1 HAs in the amino-acid preferences of many sites}
An obvious hypothesis for why the H1 deep mutational scanning is less useful for understanding the evolution of H3N2 influenza viruses is that the effect of the same mutation is often different between these two HA subtypes.
To determine if this is the case, we examined how much the amino-acid preferences of homologous sites have shifted between H3 and H1 HAs.
Prior experiments have found only modest shifts in amino-acid preferences between two variants of influenza nucleoprotein with 94\% amino-acid identity~\cite{doud2015site} and variants of HIV envelope (Env) with 86\% amino-acid identity~\cite{haddox2018mapping}.
However, the H1 and H3 HAs are far more diverged, with only 42\% amino-acid identity (Figure~\ref{fig:distance_distribution}A).
One simple way to investigate the extent of shifts in amino-acid preferences is to correlate measurements from independent deep mutational scanning replicates on H1 and H3 HAs.
Figure~\ref{fig:distance_distribution}B shows that replicate measurements on the same HA variant are more correlated than those on different HA variants.
